---
title: bench 1.0.1
date: '2018-06-28'
slug: bench-1.0.1
author: Jim Hester
categories: [package]
description: >
  bench 1.0.1 is now available on CRAN. bench allows you to benchmark
  code, tracking execution time, memory allocations and garbage
  collections.
photo:
  url: https://unsplash.com/photos/8EzNkvLQosk
  author: Maarten van den Heuvel
---

[bench] is now available on CRAN!

The goal of [bench] is to benchmark code, tracking execution time,
memory allocations and garbage collections.

Install the latest version with:
``` r
install.packages("bench")
```

### Comparison vs existing methods

Compared to other methods such as [system.time], [rbenchmark], [tictoc] or
[microbenchmark] we feel it has a number of benefits.

- Always uses the highest precision APIs available for each operating system (often nanosecond-level).
- Tracks memory allocations for each expression benchmarked.
- Tracks the number and type of R garbage collections per expression iteration.
- Verifies equality of expression results by default, to avoid accidentally
  benchmarking inequivalent code.
- `bench::press()`, a function which allows you to easily perform and combine
  benchmarks across a large grid of values.
- Uses adaptive stopping by default, running each expression for a set amount
  of time rather than for a specific number of iterations.
- Expressions are run in batches and summary statistics are calculated after
  filtering out iterations with garbage collections. This allows you to isolate
  the performance  and effects of garbage collection on running time (for more details see [Neal
  2014](https://radfordneal.https://dynverse.github.io/dynverse/wordpress.com/2014/02/02/inaccurate-results-from-microbenchmark/)).

### Usage

#### `bench::mark()`

Benchmarks can be run with `bench::mark()`, which takes one or more expressions
to benchmark against each other.

```{r example}
library(bench)
set.seed(42)
dat <- data.frame(x = runif(10000, 1, 1000), y=runif(10000, 1, 1000))
```

`bench::mark()` will throw an error if the results are not equivalent, so you
don't accidentally benchmark inequivalent code.
```{r example1-1, error = TRUE}
bench::mark(
  dat[dat$x > 500, ],
  dat[which(dat$x > 499), ],
  subset(dat, x > 500))
```

Results are easy to interpret, with human readable units.
```{r example1-2}
bnch <- bench::mark(
  dat[dat$x > 500, ],
  dat[which(dat$x > 500), ],
  subset(dat, x > 500))
bnch
```

By default the summary uses absolute measures, however relative results can be
obtained by using `relative = TRUE` in your call to `bench::mark()` or calling
`summary(relative = TRUE)` on the results.
```{r example1-3}
summary(bnch, relative = TRUE)
```

#### `bench::press()`

`bench::press()` is used to run benchmarks against a grid of parameters.
Provide setup and benchmarking code as a single unnamed argument then define
sets of values as named arguments. The full combination of values will be
expanded and the benchmarks are then _pressed_ together in the result. This
allows you to benchmark a set of expressions across a wide variety of input
sizes, perform replications and other useful tasks.

```{r example2, cache = TRUE}
set.seed(42)

create_df <- function(rows, cols) {
  as.data.frame(setNames(
    replicate(cols, runif(rows, 1, 1000), simplify = FALSE),
    rep_len(c("x", letters), cols)))
}

results <- bench::press(
  rows = c(10000, 100000),
  cols = c(10, 100),
  {
    dat <- create_df(rows, cols)
    bench::mark(
      min_iterations = 100,
      bracket = dat[dat$x > 500, ],
      which = dat[which(dat$x > 500), ],
      subset = subset(dat, x > 500)
    )
  }
)
results
```

### Plotting

`ggplot2::autoplot()` can be used to generate an informative default plot. This
plot is colored by GC level (0, 1, or 2) and faceted by parameters (if any). By
default it generates a
[beeswarm](https://github.com/eclarke/ggbeeswarm#geom_quasirandom) plot,
however you can also specify other plot types (`jitter`, `ridge`,
`boxplot`, `violin`). See `?autoplot.bench_mark` for full details.

```{r autoplot, message = FALSE, warning = FALSE, cache = TRUE, dependson = "example2"}
ggplot2::autoplot(results)
```

You can also produce fully custom plots by un-nesting the results and
working with the data directly.

```{r custom-plot, message = FALSE, cache = TRUE, dependson = "example2"}
library(tidyverse)
results %>%
  unnest() %>%
  mutate(expression = as.character(expression)) %>%
  filter(gc == "none") %>%
  ggplot(aes(x = mem_alloc, y = time, color = expression)) +
    geom_point() +
    scale_color_brewer(type = "qual", palette = 3)
```

### Dependency load

When the development version of **bench** was
[introduced](https://twitter.com/jimhester\_/status/996063591433416704) a few
users expressed concern over the number of dependencies in the package. I will
attempt to explain why these dependencies exist and why the true load may actually
be less than you might think.

While bench currently has 19 dependencies, only 8 of these are hard
dependencies; that is they are needed to install the package. Of these 8 hard
dependencies 3 of them (methods, stats, utils) are base packages installed with
R. Of these 5 remaining packages 3 have no additional dependencies (glue,
profmem, rlang). The two remaining packages (tibble and pillar) are used to
provide nice printing of the times and memory sizes and support for list
columns to store the timings, garbage collections, and allocations. These are
major features of the bench package and would not work without these
dependencies.

The remaining 11 packages are soft dependencies, used either for testing or for
optional functionality, most notably plotting.

### Feedback wanted!

We hope **bench** is a useful tool for benchmarking short expressions of code.
Please open [GitHub issues](https://github.com/r-lib/bench) for any feature
requests or bugs.

Learn more about **bench** at
  - http://bench.r-lib.org

[bench]: https://bench.r-lib.org
[rbenchmark]: https://cran.r-project.org/package=rbenchmark
[tictoc]: https://cran.r-project.org/package=tictoc
[microbenchmark]: https://cran.r-project.org/package=microbenchmark
[system.time()]: https://www.rdocumentation.org/packages/base/versions/3.5.0/topics/system.time
